Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-03 paper_jasa.tex
experiment2 setup

\begin{table*}[hbtp] 
\caption{Test Statistics for Testing Hypothesis Related to Model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
\centering 
\begin{tabular}{m{0.5cm}m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
\hline\hline 
Case & Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
\hline 
1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid. For lineup plot, we simulate data from fitted null model. \\ 
2 & $H_0: \beta_k=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.45}{\includegraphics{stat_beta_k.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid. For lineup plot, we simulate data from fitted null model. \\ 
3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\
4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid. For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space 
5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid. For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$. For lineup plot, we simulate data from standard normal. \\ 				
7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression. For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$. \\ 
8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data. For lineup plot, we simulate data from fitted model. \\[1ex] % [1ex] adds vertical space 
9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\			
\hline 
\end{tabular} 
\label{tbl:stat_multiple} 
\end{table*}

@@

This experiment is designed for the same regression model (Equation \ref{multi}) with $p=1$ and covariate $X_1$ being continuous. $X_1$ is simulated from standard normal(0,1). Data is simulated with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The different values for slope parameter that are used to generate data are shown in Table \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ while generating actual data. For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups.

@@

@@

\begin{figure}[htp]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous. One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true. Which plot is the most different from the others, in the sense that there is the steepest slope?}
       \label{fig:test_continous}
\end{figure}

>>>>

\begin{table*}[hbtp] 
\caption{Test Statistics for Testing Hypothesis Related to Model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
\centering 
\begin{tabular}{m{0.5cm}m{3cm}m{2cm}m{3cm}m{5.5cm}} 
\hline\hline 
Case & Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
\hline 
1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid. For lineup plot, we simulate data from fitted null model. \\ 
2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{3cm}   \scalebox{0.4}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Residual vs $X_k$ plots. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\
4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid. For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space 
5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid. For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$. For lineup plot, we simulate data from standard normal. \\ 				
7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression. For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$. \\ 
8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data. For lineup plot, we simulate data from fitted model. \\[1ex] % [1ex] adds vertical space 
9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\			
\hline 
\end{tabular} 
\label{tbl:stat_multiple} 
\end{table*}

@@

This experiment is designed for the same regression model (Equation \ref{multi}) with $p=1$ and covariate $X_1$ being continuous. $X_1$ is simulated from standard normal(0,1). Data is simulated with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The different values for slope parameter that are used to generate data are shown in Table \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ while generating actual data. For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups. Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots.

@@

A lineup for this experiment is shown in figure \ref{fig:test_continuous}.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11.

@@

\begin{figure}[htp]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous. One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true. Which plot is the most different from the others, in the sense that there is the steepest slope?}
       \label{fig:test_continuous}
\end{figure}
